cubble An R Package for Structuring Spatio-temporal Data

### 2022-04-23 3-advanced.Rmd
a new skeleton of intro incorporating various past writing
polish wording

# Other features and considerations

@@

One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are. This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data]. \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task. The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

>>>>

# Other features and considerations {#others}

@@

One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are. This can be seen as a matching problem [@stuart2010matching; @mcintosh2018using] that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources [@cocchi2019data]. \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task. The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}, \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

### 2022-04-23 4-1-basic.Rmd
work on section 3 the cubble package
polish wording

## Victoria covid data

This example illustrate casting different sources into a cubble object. Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA). This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics. Variable \code{covid} and \code{lga} show the first 5 rows of these two data:

@@

To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}). \code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.

@@

\pkg{cubble} will make an attempt to pair the unmatched sites as well as showing unpaired locations. These can be useful information to help clean up the data.

@@

covid <- covid %>% filter(!lga %in% pair$others$temporal)
```

After cleaning, you can cast the data into cubble with the same syntax without warnings:

```{r echo = TRUE}
cb <- as_cubble(data = list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))
```

## Australia historical maximum temperature

@@

Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station. New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:

>>>>

## Victoria covid data {#covid}

Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA). This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics. The first five rows of both data are printed below:

@@

Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create). \code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.

@@

\pkg{cubble} will attempt to pair the unmatched sites as well as those that cannot be paired. These can be helpful to clean up the data before creating the cubble again:

@@

covid <- covid %>% filter(!lga %in% pair$others$temporal)

cb <- as_cubble(data = list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))
```

## Australian historical maximum temperature

@@

Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and is followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml). The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

A few stations do not have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported. These stations are filtered out by examining whether the summarised time series has 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:

### 2022-04-23 4-2-agg.Rmd
polish wording

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level. Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map. This example shows how to organise data at both level with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data: subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

>>>>

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level. Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map. This example shows how to organise data at both levels with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.

@@

Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data. Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

### 2022-04-23 4-3-water-level.Rmd
polish wording

Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature. In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.

@@

From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station. As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020. The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}. \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if precipitation will be reflected by the water level in the river. This puts precipitation \code{prcp}, as the independent. Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.

@@

Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series. There are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.

```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected by the water level."}
knitr::include_graphics(here::here("figures/matching.png"))
```

>>>>

Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature. In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.

@@

From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station. As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020. The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}. \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation. This puts precipitation \code{prcp}, as the independent. Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.

@@

Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series. There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.

```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation."}
knitr::include_graphics(here::here("figures/matching.png"))
```

### 2022-04-23 4-5-interactive.Rmd
polish wording

---
title: "4-5-interactive"
output: html_document
---
## Interative graphic

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series. Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} objects separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month. With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo code that outlines the process to construct an interactive graphic described above:

@@

In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic. On the map, most regions in Australia have low variance of temperature range while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes. In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right. The third row the lowest temperature in August and this corresponds to Thredbo AWS in the Victoria and New South Wales border. Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.

@@

This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range can be displayed as a small subplot upon clicking on the map. This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:

@@

```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png")) 
```

>>>>

---
title: "4-5-interactive"
output: html_document
---

## Interactive graphic

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series. Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}. The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} object separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance. A ribbon band is constructed using the long form cubble to show each station's  maximum and minimum temperature across month. With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo-code that outlines the process to construct an interactive graphic described above:

@@

In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic. Most regions in Australia have low variance of temperature range On the map, while the north-west coastline, bottom of South Australia, and Victoria stand out with larger monthly changes. In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right. The third row selects the lowest temperature in August and this corresponds to Thredbo AWS on the Victoria and New South Wales border. Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.

@@

This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range is displayed as a small subplot upon clicking on the map. The procedure involves first creating the popup plots from the long form cubble as a vector and then adding these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:

@@

```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png")) 
```

### 2022-04-23 arxiv/cubble-arxiv.tex
polish wording

One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are. This can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al. 2018}) that pairs up similar similar time series in nearby locations or a data fusing exercise that merges data collected from different sources (\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}). \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task. The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al. (\protect\hyperlink{ref-Wickham2012-yr}{2012})). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

This example illustrate casting different sources into a cubble object. Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA). This data can be useful for visualising COVID spread when combined with LGA map information, which is available from Australian Bureau of Statistics. Variable \code{covid} and \code{lga} show the first 5 rows of these two data:

@@

To cast these two data into a cubble object, supply them as a list with name \code{spatial} and \code{temporal} and other arguments (\code{key}, \code{index}, and \code{coords}). \code{as_cubble()} will automatically check the match of spatial and temporal information in the two datasets and warnings will emit if a location has missing spatial or temporal information.

@@

\NormalTok{covid }\OtherTok{\textless{}{-}}\NormalTok{ covid }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{lga }\SpecialCharTok{\%in\%}\NormalTok{ pair}\SpecialCharTok{$}\NormalTok{others}\SpecialCharTok{$}\NormalTok{temporal)}
\end{Highlighting}
\end{Shaded}

After cleaning, you can cast the data into cubble with the same syntax without warnings:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{data =} \FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
                \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat))}
\end{Highlighting}
\end{Shaded}

\hypertarget{australia-historical-maximum-temperature}{%
\subsection{Australia historical maximum temperature}\label{australia-historical-maximum-temperature}}

@@

Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station. New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S) and this will be a problem when mapping more stations in the national level. Aggregation can be helpful to group series into clusters before visualising the cluster with glyph map. This example shows how to organise data at both level with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature. In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.

@@

Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series. There are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected by the water level.}\label{fig:matching}
\end{figure}

@@

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}). It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere. Readers interested in the analysis of this figure can refer to Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al. (\protect\hyperlink{ref-simmons2020global}{2020}) and Simmons et al. (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.

@@

\hypertarget{interative-graphic}{%
\subsection{Interative graphic}\label{interative-graphic}}

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series. Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} objects separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month. With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo code that outlines the process to construct an interactive graphic described above:

@@

\begin{figure}
\includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid} \caption{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
\end{figure}

>>>>

One task that may interest analysts in spatio-temporal data is to find how similar the time series from nearby sites are. This can be seen as a matching problem (\protect\hyperlink{ref-stuart2010matching}{Stuart 2010}; \protect\hyperlink{ref-mcintosh2018using}{McIntosh et al. 2018}) that pairs up similar time series in nearby locations or a data fusing exercise that merges data collected from different sources (\protect\hyperlink{ref-cocchi2019data}{Cocchi 2019}). \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task. The algorithm first matches the two data sources spatially through computing the pairwise distance on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrates this temporal matching in more details. Given two series \code{A} and \code{a}, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}, \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al. (\protect\hyperlink{ref-Wickham2012-yr}{2012})). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA). This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics. The first five rows of both data are printed below:

@@

Cubble can be created with separate spatial and temporal table in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \ref{create}. \code{as_cubble()} will automatically check the match of site in both tables emit a warning if a location has missing spatial or temporal information.

@@

\NormalTok{covid }\OtherTok{\textless{}{-}}\NormalTok{ covid }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\NormalTok{lga }\SpecialCharTok{\%in\%}\NormalTok{ pair}\SpecialCharTok{$}\NormalTok{others}\SpecialCharTok{$}\NormalTok{temporal)}

\NormalTok{cb }\OtherTok{\textless{}{-}} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{data =} \FunctionTok{list}\NormalTok{(}\AttributeTok{spatial =}\NormalTok{ lga, }\AttributeTok{temporal =}\NormalTok{ covid),}
                \AttributeTok{key =}\NormalTok{ lga, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(cent\_long, cent\_lat))}
\end{Highlighting}
\end{Shaded}

\hypertarget{australian-historical-maximum-temperature}{%
\subsection{Australian historical maximum temperature}\label{australian-historical-maximum-temperature}}

@@

Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with ``ASN00'' and is followed by the \href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of Meteorology (BOM) station number}. The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level. Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map. This example shows how to organise data at both levels with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for 639 stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.

@@

Bureau of Meteorology collects \href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water data} from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature. In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.

@@

Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series. There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation.}\label{fig:matching}
\end{figure}

@@

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}). It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere. Readers interested in the analysis of this figure can refer to Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al. (\protect\hyperlink{ref-simmons2020global}{2020}), and Simmons et al. (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.

@@

\hypertarget{interactive-graphic}{%
\subsection{Interactive graphic}\label{interactive-graphic}}

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns, such as trend or seasonality, in the time series. Combining these two types of plots with interactivity lets users link between points on the map and the corresponding time series to explore the spatial and temporal dimensions of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}. The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} object separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance. A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature across month. With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo-code that outlines the process to construct an interactive graphic described above:

@@

{\centering \includegraphics[width=0.45\linewidth,height=0.25\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-mid}

\caption{Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.}\label{fig:interactive-popup}
\end{figure}
